2.3 Prob Model for PCR Positivity Estimation

### 2024-03-11 2_scripts/2.1.1 Canpath Data.qmd
Exp: model based on samples from traning data

## recode QC as missing
  m2$region[m2$region=="QC"]<-NA # N=4 coded as missing (NA)
```
## see what is the assay used in each region
```{r}
m2 %>%
  select(region, assay_id) %>%
  group_by(region, assay_id) %>%
  summarise(Freq=n())
```
## save the data
```{r}
Canpaths<-m2
rm(m1, m2, n_admin, ser_sw)
save(Canpaths, file = "../1_data/private/CANPATH/Canpaths.RData")
```

>>>>

## recode QC as missing
  m2$region[m2$region=="QC"]<-NA # N=4 coded as missing (NA)
```
## see what is the assay used in each region
```{r}
m2 %>%
  select(region, assay_id) %>%
  group_by(region, assay_id) %>%
  summarise(Freq=n())

# clean the data var
m2$date_col<-as.Date(m2$date_col, format = "%Y-%m-%d")
m2$date_asy<-as.Date(m2$date_asy, format = "%Y-%m-%d")

# when those assay was applied?
m2 %>%
  group_by(region) %>%
  summarise(Freq=n(),
            tcol_min = min(date_col, na.rm=T),
            tcol_max = max(date_col, na.rm=T))
```
## save the data
```{r}
Canpaths<-m2
rm(m1, m2, n_admin, ser_sw)
save(Canpaths, file = "../1_data/private/CANPATH/Canpaths.RData")
```

### 2024-03-11 2_scripts/2.3 Prob Model for PCR pos.qmd
Exp: model based on samples from traning data

##============ Part III. Apply Model to CBS Data ================

@@

## Part IV Apply model to APL data

@@

## Part V Apply model to CANPATH data

@@

table(Canpath$C1_CITF_ASSAY_ID)

```

>>>>

##============ Part III-1 Apply Model to CBS Data ================

@@

## Part III-2 Apply model to APL data

@@

## Part III-3 Apply model to CANPATH data

@@

table(Canpath$C1_CITF_ASSAY_ID)
```

## Part IV Misc Analysis

1. About the age dist in training data and CBS: The difference on age distribution might be a potential problem. Generally speaking, blood donors (CBS, FR-Leon data) is relatively older, while Covid-19 convalescent patients are younger.
```{r}
# The FR-Leon data
summary(dta_all[dta_all$study==1,]$age) # mean=46, median = 48
hist(dta_all[dta_all$study==1,]$age)

# UK - Convalscent Pts
summary(dta_all[dta_all$study==6,]$age) # mean=44.54, median = 47
hist(dta_all[dta_all$study==6,]$age, breaks = 25) # two waves?

# overall dist (FR-Leon + UK)
summary(dta_all[dta_all$study %in% c(1,6),]$age) # mean=46, median = 48
hist(dta_all[dta_all$study %in% c(1,6),]$age, breaks = 25)
# overall it looks good

# CBS
summary(CBSs$age) # one ourlier = 168, who is this??
hist(CBSs$age[CBSs$age<100])

# CBS by anti-N pos
hist(CBSs$age[CBSs$age<100 & 
              CBSs$roche_n>1])

# KS test: UK+FR-Leon vs CBS
ks.test(dta_all[dta_all$study %in% c(1,6),]$age, 
        CBSs$age[CBSs$age<100])
# as KS test suggests, the distribution of age btw two groups is different
```
```{r}
# create a sampling prob variable, based on the age dist in CBS
CBSs$age_group<-CBS$age_group

tb1<-as.data.frame(prop.table(table(CBS$age_group)))
colnames(tb1)<-c("age_group","age_w")
tb1$n_sp<-round(tb1$age_w*650, digits = 0)

dta_all$age_group[dta_all$age<20 & dta_all$age>0]<-"10-19 years"
dta_all$age_group[dta_all$age<30 & dta_all$age>20]<-"20-29 years"
dta_all$age_group[dta_all$age<40 & dta_all$age>30]<-"30-39 years"
dta_all$age_group[dta_all$age<50 & dta_all$age>40]<-"40-49 years"
dta_all$age_group[dta_all$age<60 & dta_all$age>50]<-"50-59 years"
dta_all$age_group[dta_all$age<70 & dta_all$age>60]<-"60-69 years"
dta_all$age_group[dta_all$age<80 & dta_all$age>70]<-"70-79 years"
dta_all$age_group[dta_all$age>80]<-"80+ years"
table(dta_all$age_group, useNA = 'ifany')

dta_all<-merge(dta_all, tb1, by="age_group", all.x=T, all.y=F)
```

#### Create a sample based on the sampling prob:
```{r}
# a simple prob sampling:
set.seed(2048)
dta_sp<-slice_sample(dta_all[!is.na(dta_all$age_w),], n=650,
             weight_by = dta_all[!is.na(dta_all$age_w),]$age_w,
             replace = FALSE)

# how does age looks like in the sample?
hist(dta_sp$age)

# Stratified sampling (Alternative):
dta1<-dta_all[!is.na(dta_all$age_group == "10-19 years"),] %>%
  slice_sample(n=tb1$n_sp[2], replace = FALSE)
dta2<-dta_all[!is.na(dta_all$age_group == "20-29 years"),] %>%
  slice_sample(n=tb1$n_sp[3], replace = FALSE)
dta3<-dta_all[!is.na(dta_all$age_group == "30-39 years"),] %>%
  slice_sample(n=tb1$n_sp[4], replace = FALSE)
dta4<-dta_all[!is.na(dta_all$age_group == "40-49 years"),] %>%
  slice_sample(n=tb1$n_sp[5], replace = FALSE)
dta5<-dta_all[!is.na(dta_all$age_group == "50-59 years"),] %>%
  slice_sample(n=tb1$n_sp[6], replace = FALSE)
dta6<-dta_all[!is.na(dta_all$age_group == "60-69 years"),] %>%
  slice_sample(n=tb1$n_sp[7], replace = FALSE)
dta7<-dta_all[!is.na(dta_all$age_group == "70-79 years"),] %>%
  slice_sample(n=tb1$n_sp[8], replace = FALSE)
dta8<-dta_all[!is.na(dta_all$age_group == "80+ years"),] %>%
  slice_sample(n=sum(tb1$n_sp[9:10]), replace = FALSE)
dta_sp<-rbind.data.frame(dta1, dta2, dta3, dta4, dta5,
                         dta6, dta7, dta8)
rm(dta1, dta2, dta3, dta4, dta5, dta6, dta7, dta8)

# KS test
ks.test(dta_sp$age, CBSs$age[CBSs$age<100])
```

## Regression on sampled data:
```{r}
mnlm03.s<-multinom(relevel(as.factor(time_pso), ref = "never_inf") ~ log_roche_n + 
                     age + gender, 
                 data = dta_sp,
                 weights = sample_wt)
summary(mnlm03.s)
```
